--------------------------------------------------------------------------------
title: "Pharmacy First — Practice-level data — Medication Analysis"
author: "Arnaud Iradukunda"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
--------------------------------------------------------------------------------
# 1. Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(nlme)
library(geepack)
library(scales)
library(readr)
library(here)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)
```
# 2. Data and intervention dates
```{r data}

practice_level_data <- read_csv(
  here("output", "practice_level_data_by_practice_only.csv")
)

GP_connect <- as.Date("2025-10-01")
PF_start   <- as.Date("2024-02-01")

practice_level_data <- practice_level_data %>%
  clean_names() %>%  # This comes from the janitor package and standardises column names (no spaces).
  mutate(interval_start = as.Date(interval_start)
         ) 

head(practice_level_data)
names(practice_level_data)

conditions <- c(
  "uti",
  "sinusitis",
  "insectbite",
  "otitismedia",
  "sorethroat",
  "shingles",
  "impetigo",
  "all_conditions"
)

control_map <- c(
  acutebronchitis_control         = "Acute bronchitis (control)",
  conjunctivitisallergic_control  = "Allergic conjunctivitis (control)",
  vulvovaginalcandidiasis_control = "Vulvovaginal candidiasis (control)"
)

head(practice_level_data)
#names(practice_level_data)
```

# 3. Monthly totals across practices
The dataset is aggregated to monthly totals across practices for GP and Pharmacy First (PF) medication counts.
```{r monthly-long}
monthly <- practice_level_data %>%
  group_by(interval_start) %>%
  summarise(
    population = sum(population, na.rm = TRUE),
    # Eligible populations
    #Conditions specific
    elig_uti         = sum(populationeligible_uuti, na.rm = TRUE),
    elig_sinusitis   = sum(populationeligible_sinusitis, na.rm = TRUE),
    elig_insectbite  = sum(populationeligible_insectbite, na.rm = TRUE),
    elig_otitismedia = sum(populationeligible_otitismedia, na.rm = TRUE),
    elig_sorethroat  = sum(populationeligible_sorethroat, na.rm = TRUE),
    elig_shingles    = sum(populationeligible_shingles, na.rm = TRUE),
    elig_impetigo    = sum(populationeligible_impetigo, na.rm = TRUE),
    # Controls specific
    elig_acutebronchitis_control = sum(populationeligible_acutebronchitis_control, na.rm = TRUE),
    elig_vulvovaginalcandidiasis_control= sum(populationeligible_vulvovaginalcandidiasis_control, na.rm = TRUE),
    elig_conjunctivitisallergic_control = sum(populationeligible_conjunctivitisallergic_control, na.rm = TRUE),
    across(
      starts_with("gp_medication_"),
      ~ sum(.x, na.rm = TRUE)   # Find all variables whose names start with gp_medication_ and sum each one within each month.
    ),
    across(
      starts_with("pf_medication_"),
      ~ sum(.x, na.rm = TRUE)  #  Find all variables whose names start with pf_medication_ and sum each one within each month (.x represents the current column being processed).
    ),
    .groups = "drop"          #  This remove the grouping (our dataset is now a normal data frame)
  ) %>%
  arrange(interval_start) %>%
  pivot_longer(                                               
    cols = starts_with(c("gp_medication_", "pf_medication_")), # GP/PF medication (wide to long)
    names_to = c("Setting", "condition"),                      # Setting = gp or Setting = pf, condition = uti

    names_pattern = "^(gp|pf)_medication_(.*)$",             #
    values_to = "medication_count"
  ) %>%
  mutate(
    Setting = toupper(Setting),
    time_index = match(
      interval_start,
      sort(unique(interval_start))
    ) - 1,
    post = as.integer(interval_start >= GP_connect),
    time_since_int = if_else(
      post == 1,
      time_index - min(time_index[post == 1]),
      0
    )
  ) %>%
  arrange(interval_start, Setting, condition)
names(monthly)
```
# 4. Monthly_long ( this pivot eligible pop into long-format)
```{r monthly_long}
library(dplyr)
library(tidyr)
monthly_long <- monthly %>%
  pivot_longer(
    cols = starts_with("elig_"),
    names_to = "eligible_condition",
    values_to = "eligible_population"
  ) %>%
  mutate(
    condition = case_when(
      eligible_condition == "elig_uti" ~ "uti",
      eligible_condition == "elig_sinusitis" ~ "sinusitis",
      eligible_condition == "elig_insectbite" ~ "insectbite",
      eligible_condition == "elig_otitismedia" ~ "otitismedia",
      eligible_condition == "elig_sorethroat" ~ "sorethroat",
      eligible_condition == "elig_shingles" ~ "shingles",
      eligible_condition == "elig_impetigo" ~ "impetigo",
      eligible_condition == "elig_acutebronchitis_control" ~ "acutebronchitis_control",
      eligible_condition == "elig_vulvovaginalcandidiasis_control" ~ "vulvovaginalcandidiasis_control",
      eligible_condition == "elig_conjunctivitisallergic_control" ~ "conjunctivitisallergic_control"
    )
  ) %>%
  select(
    interval_start,
    population,
    Setting,
    condition,
    eligible_population,
    medication_count,
    time_index,
    post,
    time_since_int
  )
names(monthly_long)
table(monthly_long$condition)
table(monthly_long$eligible_population)
```

## 5. Descriptive analysis
## 5.1 Medication numerator over time

```{r}
table(monthly_long$condition)
```


```{r descriptive}
descr_numerator <- monthly_long %>%              # Here i can also use monthly long
  filter(condition %in% conditions) %>%
  select(
    interval_start,
    Setting,
    condition,
    medication_count
  ) %>%
  mutate(
    condition = str_to_sentence(condition)
  )
```
#

```{r}
table(descr_numerator$Setting)
```
# Graph 1: Medication counts
```{r graph_A1, fig.width=10, fig.height=20, fig.cap="Figure 1. Medication counts by condition and setting (GP vs Pharmacy First)."}
graph_A1_numerator <- ggplot(
  descr_numerator,aes(x = interval_start, y = medication_count,colour = Setting)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(xintercept = as.numeric(GP_connect),linetype = "dashed",colour = "black") +
  facet_wrap(~ condition,scales = "free_y",ncol = 1) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 months") +
  #facet_wrap(~ condition) +
  scale_colour_manual(
    values = c(
      "GP" = "steelblue",
      "PF" = "firebrick"
    )
  ) +
  labs(
    title = "Medication counts by condition and setting",
    subtitle = "Dashed line = GP Connect",
    x = NULL,
    y = "Number of medications (numerator)",
    colour = "Setting"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
graph_A1_numerator

# Save as PNG
ggsave(
  filename = "output/results_3_Fig1_medication_counts_by_condition.png",
  plot = graph_A1_numerator,
  path = here::here(),
  width = 10,
  height = 20,
  dpi = 300
)
```

## 6.2 Medication rate over time

```{r descriptive-numerator}
descr_rate <- monthly_long %>%
  filter(condition %in% conditions) %>%
  select(
    interval_start,
    Setting,
    condition,
    eligible_population,
    medication_count
  ) %>%
  mutate(
    condition = str_to_sentence(condition), # This capitalises the first letter and makes the rest lowercase.
    rate_per_1000 = (medication_count/eligible_population)*1000
  )
```

#graph 2: Medication rate
```{r graph_A2_rate, fig.width=10, fig.height=20, fig.cap="Figure 2. Medication rate per 1,000 population by condition and setting."}
graph_A2_rate <- ggplot(descr_rate,aes(x = interval_start,y = rate_per_1000,colour = Setting)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline( xintercept = as.numeric(GP_connect),linetype = "dashed",colour = "black") +
  facet_wrap( ~ condition,scales = "free_y",ncol = 1) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 months") +
  #facet_wrap(~ condition) +
  scale_colour_manual(
    values = c(
      "GP" = "steelblue",
      "PF" = "darkred"
    )
  ) +
  labs(
    title = "Medication rate by condition and setting",
    subtitle = "Rate per 1,000 population; dashed line = GP Connect",
    x = NULL,
    y = "Rate per 1,000 population",
    colour = "Setting"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
graph_A2_rate

# Save as PNG
ggsave(
  filename = "output/results_3_Fig2_medication_rate_by_condition.png",
  plot = graph_A2_rate,
  path = here::here(),
  width = 10,
  height = 20,
  dpi = 300
)
```
# 7. GP intervention-control descriptive plots
```{r intervention-control-data}
descr_pairs <- practice_level_data %>%
  select(
    interval_start,
    gp_medication_uti,
    gp_medication_vulvovaginalcandidiasis_control,
    gp_medication_sorethroat,
    gp_medication_acutebronchitis_control,
    gp_medication_insectbite,
    gp_medication_conjunctivitisallergic_control
  ) %>%
  group_by(interval_start) %>%
  mutate(
    medication_uti =
      sum(gp_medication_uti, na.rm = TRUE),

    medication_vulvovaginalcandidiasis_control =
      sum(gp_medication_vulvovaginalcandidiasis_control, na.rm = TRUE),

    medication_sorethroat =
      sum(gp_medication_sorethroat, na.rm = TRUE),

    medication_acutebronchitis_control =
      sum(gp_medication_acutebronchitis_control, na.rm = TRUE),

    medication_insectbite =
      sum(gp_medication_insectbite, na.rm = TRUE),

    medication_conjunctivitisallergic_control =
      sum(gp_medication_conjunctivitisallergic_control, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  select(
    interval_start,
    starts_with("medication_")
  ) %>%
  distinct() %>%
  pivot_longer(
    cols = starts_with("medication_"),
    names_to = "condition",
    values_to = "count"
  ) %>%
  mutate(
    condition = recode(
      condition,
      medication_uti =
        "UTI",
      medication_vulvovaginalcandidiasis_control =
        "Vulvovaginal candidiasis",
      medication_sorethroat =
        "Sore throat",
      medication_acutebronchitis_control =
        "Acute bronchitis",
      medication_insectbite =
        "Insect bite",
      medication_conjunctivitisallergic_control =
        "Allergic conjunctivitis"
    )
  )
```

## 7.1 UTI vs vulvovaginal candidiasis
```{r descr_pairs}
table(descr_pairs$condition)
```
## Graph
```{r graph-uticontrol, fig.width=10, fig.height=5, fig.cap="Figure 3. GP medications for UTI and vulvovaginal candidiasis (control)."}
graph_A3_uti <- descr_pairs %>%
  filter(
    condition %in%
      c(
        "UTI",
        "Vulvovaginal candidiasis"
      )
  ) %>%
  ggplot(aes(x = interval_start,y = count,colour = condition)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline( xintercept = as.numeric(GP_connect),linetype = "dotted") +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 months") +
  #facet_wrap(~ condition) +
  labs(
    title = "GP medications for UTI and Vulvovaginal candidiasis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Number of medications (numerator)",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_A3_uti

# Save as PNG
ggsave(
  filename = "output/results_3_Fig3_uti_vs_vulvovaginal_candidiasis.png",
  plot = graph_A3_uti,
  path = here::here(),
  width = 10,
  height = 5,
  dpi = 300
)
```

## 7.2 Sore throat vs acute bronchitis

```{r graph-sorethroat-control, fig.width=10, fig.height=5, fig.cap="Figure 4. GP medications for sore throat and acute bronchitis (control)."}
graph_A3_sorethroat <- descr_pairs %>%
  filter(
    condition %in%
      c(
        "Sore throat",
        "Acute bronchitis"
      )
  ) %>%
  ggplot(aes(x = interval_start,y = count,colour = condition)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(xintercept = as.numeric(GP_connect),linetype = "dotted") +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 months") +
  #facet_wrap(~ condition) +
  labs(
    title = "GP medications for sore throat and acute bronchitis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Number of medications (numerator)",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_A3_sorethroat

# Save as PNG
ggsave(
  filename = "output/results_3_Fig4_sorethroat_vs_acutebronchitis.png",
  plot = graph_A3_sorethroat,
  path = here::here(),
  width = 10,
  height = 5,
  dpi = 300
)
```

## 7.3 Insect bite vs allergic conjunctivitis

```{r graph-insectbite-control, fig.width=10, fig.height=5, fig.cap="Figure 5. GP medications for insect bite and allergic conjunctivitis (control)."}
graph_A3_insectbite <- descr_pairs %>%
  filter(
    condition %in%
      c(
        "Insect bite",
        "Allergic conjunctivitis"
      )
  ) %>%
  ggplot(
    aes(
      x = interval_start,
      y = count,
      colour = condition
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 months") +
  #facet_wrap(~ condition) +
  labs(
    title = "GP medications for insect bite and allergic conjunctivitis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Number of medications (numerator)",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_A3_insectbite

# Save as PNG
ggsave(
  filename = "output/results_3_Fig5_insectbite_vs_allergicconjunctivitis.png",
  plot = graph_A3_insectbite,
  path = here::here(),
  width = 10,
  height = 5,
  dpi = 300
)
```



#Pair rate

#Descriptives pairs rate
```{r descr_pairs_rate}
descr_pairs_rate <- monthly_long %>%
  select(
    interval_start,
    Setting,
    condition,
    eligible_population,
    medication_count
  ) %>%
  mutate(
    condition = str_to_sentence(condition), # This capitalises the first letter and makes the rest lowercase.
    rate_per_1000 = (medication_count/eligible_population)*1000
  )
```

## 8.1 UTI vs vulvovaginal candidiasis rate
```{r}
table(descr_pairs_rate$condition)
```


```{r graph-uti-control, fig.width=10, fig.height=5, fig.cap="Figure 6. GP medication rate for UTI and vulvovaginal candidiasis (control)."}
graph_B3_uti <- descr_pairs_rate %>%
  filter(
    condition %in%
      c(
        "Uti",
        "Conjunctivitisallergic_control"
      )
  ) %>%
  ggplot(
    aes(
      x = interval_start,
      y = rate_per_1000,
      colour = condition
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 months") +
  #facet_wrap(~ condition) +
  labs(
    title = "GP medications for UTI and Vulvovaginal candidiasis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Rate per 1,000 population",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_B3_uti

# Save as PNG
ggsave(
  filename = "output/results_3_Fig6_rate_uti_vs_vulvovaginal_candidiasis.png",
  plot = graph_B3_uti,
  path = here::here(),
  width = 10,
  height = 5,
  dpi = 300
)
```

## 8.2 Sore throat vs acute bronchitis rate
```{r graph-sorethroatcontrol, fig.width=10, fig.height=5, fig.cap="Figure 7. GP medication rate for sore throat and acute bronchitis (control)."}
graph_B3_sorethroat <- descr_pairs_rate %>%
  filter(
    condition %in%
      c(
        "Sorethroat",
        "Acutebronchitis_control"
      )
  ) %>%
  ggplot(
    aes(
      x = interval_start,
      y = rate_per_1000,
      colour = condition
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 months") +
  #facet_wrap(~ condition) +
  labs(
    title = "GP medications for sore throat and acute bronchitis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Rate per 1000 population",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_B3_sorethroat

# Save as PNG
ggsave(
  filename = "output/results_3_Fig7_rate_sorethroat_vs_acutebronchitis.png",
  plot = graph_B3_sorethroat,
  path = here::here(),
  width = 10,
  height = 5,
  dpi = 300
)
```

## 8.3 Insect bite vs allergic conjunctivitis rate
```{r graph-insectbitecontrol, fig.width=10, fig.height=5, fig.cap="Figure 8. GP medication rate for insect bite and allergic conjunctivitis (control)."}
graph_B3_insectbite <- descr_pairs_rate %>%
  filter(
    condition %in%
      c(
        "Insectbite",
        "Conjunctivitisallergic_control"
      )
  ) %>%
  ggplot(aes(x = interval_start,y = rate_per_1000,colour = condition)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  geom_vline(
    xintercept = as.numeric(GP_connect),
    linetype = "dotted"
  ) +
  scale_x_date(
    date_labels = "%b %y",
    date_breaks = "1 month"
  ) +
  labs(
    title = "GP medications for insect bite and allergic conjunctivitis",
    subtitle = "Dotted line = GP Connect",
    x = NULL,
    y = "Rate per 1,000 population",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom"
  )

graph_B3_insectbite

# Save as PNG
ggsave(
  filename = "output/results_3_Fig8_rate_insectbite_vs_allergicconjunctivitis.png",
  plot = graph_B3_insectbite,
  path = here::here(),
  width = 10,
  height = 5,
  dpi = 300
)
```

# 8. ITSA models — GP
The following section prepares an interrupted time-series analysis for all seven Pharmacy First conditions.
The intervention is **PF service implementation on 1 february 2024**.
However,**GP Connect on 1 October 2025** may influence reported data in GP records.

We model the monthly GP-recorded medication counts using a segmented
Poisson regression with an offset for the population at risk.

The model specification is:

\[
\log\left(E(Y_{gt})\right)
=
\beta_0
+
\beta_1 T_{gt}
+
\beta_2 X_{gt}
+
\beta_3 X_{gt}T_{gt}
+
\log(Pop_{gt})
\tag{7}
\]

where:

- \(Y_{gt}\) is the observed count of the outcome for practice \(g\) at month \(t\);
- \(E(Y_{gt})\) is the expected count;
- \(T_{gt}\) is the time in months from the start of the study period;
- \(X_{gt}\) is an indicator for the post-intervention period
  (\(1\) = at/after PF service implementation, \(0\) = before);
- \(X_{gt}T_{gt}\) is the interaction term allowing the slope to change after the intervention;
- \(Pop_{gt}\) is the population at risk for practice \(g\) at month \(t\), included as an offset.

- `β0` = baseline level of the outcome at the start of the study period;
- `β1` = baseline time trend (change in the outcome per month before PF service);
- `β2` = immediate level change in the outcome following PF service implementation;
- `β3` = change in the time trend (slope) following PF service implementation.


## 8.1 Prepare ITSA dataset

```{r}
#names(practice_level_data)
# GP Connect intervention date
intervention_date <- as.Date("2025-10-01")
#ITSA dataset

practice_level_data_itsa <- practice_level_data %>%
  mutate( interval_start = as.Date(interval_start)) %>%
  arrange(practice,interval_start) %>%
  group_by(practice) %>%
  mutate(
    # Time starts at 1 for each practice
    T = row_number()-1,
    # Intervention start
    X = as.integer(interval_start >= intervention_date),
    # Time point of intervention
    first_post_T = T[
      interval_start == intervention_date
    ][1],

    # Change in slope after  intervention
    XT = pmax(
      T - first_post_T,
      0
    ) * X
  ) %>%
  ungroup() %>%
  select(-first_post_T)

# Check data
print(head(practice_level_data_itsa, 14))

# Check ITSA variables
print(
  practice_level_data_itsa %>%
    distinct(interval_start, T, X, XT) %>%
    arrange(interval_start)
)
# Export into csv
#write_csv(practice_level_data_itsa,"practice_level_da
```
## II.GEE models:

#All Data
```{r All_Data}
practice_level_data_itsa
```

## II.1.GP UTI medication prescribing

```{r gee_uti_data}
# 1. Load packages
library(dplyr)
library(ggplot2)
library(lubridate)
library(geepack)
library(broom)
gee_uti_data <- practice_level_data_itsa %>%
  select(
    practice,
    interval_start,
    population,
    populationeligible_uuti,
    gp_medication_uti,T,
    X,
    XT
  )
gee_uti_data
```

##3. Data checks
```{r data-check}
glimpse(gee_uti_data)
summary(gee_uti_data$gp_medication_uti)
summary(gee_uti_data$populationeligible_uuti)
table(gee_uti_data$X)
range(gee_uti_data$interval_start, na.rm = TRUE)
```
# 4.Missing and invalid values
```{r missingness}
 gee_uti_data %>%
  summarise(
    n_rows = n(),
    n_practices = n_distinct(practice),
    n_na = sum(is.na(gp_medication_uti)),
    n_nan = sum(is.nan(gp_medication_uti)),
    n_inf = sum(is.infinite(gp_medication_uti)),
    min = min(gp_medication_uti, na.rm = TRUE),
    max = max(gp_medication_uti, na.rm = TRUE),
    missing_outcome = sum(is.na(gp_medication_uti)),
    missing_population = sum(is.na(population)),
    missing_eligible_population = sum(is.na(populationeligible_uuti)),
    zero_population = sum(population <= 0, na.rm = TRUE),
    zero_eligible_population = sum(populationeligible_uuti <= 0, na.rm = TRUE),
    missing_date = sum(is.na(interval_start))
  )
```
## 5. Fit the GEE model
```{r }
class(gee_uti_data$population)
names(gee_uti_data$population)
```

##gee_uti_model

```{r gee_uti_model}
gee_uti_model <- geeglm(
  gp_medication_uti ~ T + X + XT,
  family = poisson(link = "log"),
  offset = log(population), # I used pop because the eligible pop contains zero
  id = practice,
  corstr = "ar1",
  data = gee_uti_data
)
```

## 6. Model summary
```{r model_summary}
summary(gee_uti_model)
```

##7.Fitted values from the GEE model

```{r plot-gee-uti-post-trend, fig.width=10, fig.height=6, fig.cap="Figure 9. GP medications for UTI — observed values and GEE fitted trend."}
gee_uti_plot_data <- gee_uti_data %>%
  mutate(
    fitted_rate = predict( gee_uti_model,type = "response"),fitted_count = fitted_rate 
    )

gee_uti_trend_plot <- ggplot(gee_uti_plot_data,aes(x = interval_start)) +
   # Observed monthly medications
  geom_point(aes(y = gp_medication_uti), size = 1.5, alpha = 0.6
    ) +
  # Observed line
  geom_line(aes(y = gp_medication_uti),linewidth = 0.6, alpha = 0.5) +
  # GEE fitted trend
  geom_line(aes(y = fitted_count),linewidth = 1) +
  # GP Connect intervention
  geom_vline(xintercept = as.numeric(GP_connect),linetype = "dotted",linewidth = 0.8) +
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "1 months") +
  labs(
    title = "GP medications for UTI",
    subtitle = "Observed values and GEE fitted trend",
    x = NULL,
    y = "Number of medications"
  ) +
  
  theme_minimal(base_size = 11) +
  
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
gee_uti_trend_plot

# Save as PNG
ggsave(
  filename = "output/results_3_Fig9_gee_uti_observed_vs_fitted.png",
  plot = gee_uti_trend_plot,
  path = here::here(),
  width = 10,
  height = 6,
  dpi = 300
)
```


# 7. Incidence-rate ratios (IRRs)

```{r irr}
irr_results <- broom::tidy(
  gee_uti_model,
  exponentiate = TRUE,
  conf.int = TRUE
)

irr_results
```
# 8. Extract the main Pharmacy First effects
```{r pf_effects}
pf_effects <- irr_results %>%
  filter(term %in% c( "T","X", "XT")) %>%
  select(
    term,
    estimate,
    conf.low,
    conf.high,
    p.value
  )

pf_effects
```
# 9.Table
```{r pf_effects_table}
pf_effects %>%
  mutate(
    IRR = round(estimate, 3),
    CI = paste0(
      round(conf.low, 3),
      "–",
      round(conf.high, 3)
    ),
    p_value = signif(p.value, 3)
  ) %>%
  select(
    term,
    IRR,
    CI,
    p_value
  )
```
# 10. Observed monthly GP UTI medication rate
```{r observed_rate}
monthly_uti <- gee_uti_data %>%
  group_by(interval_start) %>%
  summarise(
    gp_medication_uti = sum(
      gp_medication_uti,
      na.rm = TRUE
    ),

    population = sum(
      population,
      na.rm = TRUE
    ),

    rate_per_1000 = (
      gp_medication_uti / population
    ) * 1000,

    .groups = "drop"
  )
```
# 11. Visualise observed prescribing rate
```{r observed_rate_plot, fig.width=10, fig.height=6, fig.cap="Figure 10. GP UTI antimicrobial prescribing — monthly rate before and after Pharmacy First implementation."}
observed_rate_plot <- ggplot( monthly_uti, aes(x = interval_start, y = rate_per_1000)) +
   geom_line(linewidth = 0.8) +
   geom_point(size = 1.5) +
   # Pharmacy First implementation
  geom_vline( xintercept = as.numeric(as.Date("2025-10-01")),
    linetype = "dashed",
    linewidth = 0.8 
    ) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 months") +
  annotate(
    "text",
    x = as.Date("2025-10-01"),
    y = Inf,
    label = "Pharmacy First",
    vjust = 1.5,
    hjust = -0.05
  ) +

  labs(
    title = "GP UTI antimicrobial prescribing",
    subtitle = "Monthly prescribing rate before and after Pharmacy First implementation",
    x = "Month",
    y = "GP UTI medication prescriptions per 1,000 population"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

observed_rate_plot

# Save as PNG
ggsave(
  filename = "output/results_3_Fig10_gp_uti_prescribing_rate.png",
  plot = observed_rate_plot,
  path = here::here(),
  width = 10,
  height = 6,
  dpi = 300
)
```

## 12.Fitted/predicted trend from the GEE model
```{r predicted_trend}
gee_uti <- gee_uti_data %>%
  mutate(
    predicted_rate_per_1000 =
      predict(
        gee_uti_model,
        type = "response"
      ) /
      population * 1000
  )
```

#II.Controlled ITSA
```{r }

names(practice_level_data)
```

#data-uti_and_control

```{r data-uti_and_control}
practice_level_uti_and_control_long <- practice_level_data %>%
  select(
    practice,
    interval_start,
    population,
    populationeligible_uuti,
    gp_medication_uti,
    gp_medication_vulvovaginalcandidiasis_control
     )%>%
  pivot_longer(
    cols = c(
      gp_medication_uti,
      gp_medication_vulvovaginalcandidiasis_control
    ),
    names_to = "condition",
    values_to = "medication_count"
  ) %>%
  mutate( interval_start = as.Date(interval_start)) %>%
  arrange(practice,interval_start) %>%
  group_by(practice,condition) %>% # :By both practice.condition because each       practice now has two rows per month:
                                    #         —one for UTI and one for control
                                    #         —T will increase twice per month. 
  mutate(
    # Time starts at 1 for each practice
    T = row_number()-1,
    # Intervention start
    X = as.integer(interval_start >= intervention_date),
    # Time point of intervention
    first_post_T = T[interval_start == intervention_date][1],
    # Change in slope after  intervention
    XT = pmax(T - first_post_T, 0 ) * X ) %>%
    # Multiple-group ITS interaction terms
    # Z = 1 for UTI, 0 for control
    mutate( Z = as.integer(condition == "gp_medication_uti"),
    ZT  = Z * T,
    ZX  = Z * X,
    ZXT = Z * XT
    ) %>%
    ungroup() %>%
  select(-first_post_T)
print(
  practice_level_uti_and_control_long %>%
    select(practice,condition,interval_start,T,X,XT,Z,ZT,ZX, ZXT)
)
```

#ITSA controlled model
#gee_uti_model

```{r gee_uti_controlled_model}
gee_uti_controlled_model <- geeglm(
  medication_count ~ T + X + XT + ZT + ZX + ZXT ,
  family = poisson(link = "log"),
  offset = log(population), # I used pop because the eligible pop contains zero
  id = practice,
  corstr = "ar1",
  data = practice_level_uti_and_control_long
)
summary(gee_uti_controlled_model)
exp(gee_uti_controlled_model$coefficients)
```

#2.Coefficient table with exponentiated rate ratios + CIs

```{r Coefficient_table}

broom::tidy(gee_uti_controlled_model, exponentiate = TRUE, conf.int = TRUE)

```


#2. Post-intervention trend comparisons


```{r Post-intervention-trend-comparison}

lincom <- function(model, terms, level = 0.95) {
  b <- coef(model)
  V <- vcov(model)
  cv <- setNames(rep(0, length(b)), names(b))
  cv[terms] <- 1

  est <- sum(cv * b)
  se  <- sqrt(t(cv) %*% V %*% cv)[1, 1]
  z   <- est / se
  p   <- 2 * pnorm(-abs(z))
  ci  <- est + c(-1, 1) * qnorm(1 - (1 - level) / 2) * se

  data.frame(Estimate = est, SE = se, z = z, `Pr(>|z|)` = p,
             CI.lower = ci[1], CI.upper = ci[2],
             RateRatio = exp(est), RR.lower = exp(ci[1]), RR.upper = exp(ci[2]),
             check.names = FALSE)
}

trend_results <- rbind(
  Treatment    = lincom(gee_uti_controlled_model, c("T", "ZT", "XT", "ZXT")),
  Control      = lincom(gee_uti_controlled_model, c("T", "XT")),
  Difference = lincom(gee_uti_controlled_model, c("ZT", "ZXT"))
)
trend_results
```

#post_estimation_trends-plot

```{r post_estimation_trends-plot, fig.width=10, fig.height=6, fig.cap="Figure 11. Post-estimation trends: UTI (treatment) vs vulvovaginal candidiasis (control)."}

library(dplyr)
library(ggplot2)

plot_data <- practice_level_uti_and_control_long %>%
  mutate(fitted_rate = predict(gee_uti_controlled_model, type = "response") / population) %>%
  group_by(interval_start, Z) %>%
  summarise(
    actual_rate    = sum(medication_count) / sum(population),
    predicted_rate = weighted.mean(fitted_rate, population),
    .groups = "drop"
  ) %>%
  mutate(group_label = if_else(Z == 1, "Treatment", "Control"))

intervention_date <- as.Date("2025-10-01")  #

post_estimation_trend_plot <- ggplot(plot_data, aes(x = interval_start, colour = group_label, group = group_label)) +
  geom_point(aes(y = actual_rate, shape = group_label), size = 2) +
  geom_line(aes(y = predicted_rate, linetype = group_label)) +
  geom_vline(xintercept = as.numeric(intervention_date), linetype = "dotted") +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 months") +
  scale_shape_manual(values = c(Treatment = 16, Control = 1)) +
  scale_linetype_manual(values = c(Treatment = "solid", Control = "dashed")) +
  labs(
    title = "Urinary tract infections (Treatment) vs. Vulvovaginalis-candidiasis (Control)",
    subtitle = paste("GP connect starts:", format(intervention_date, "%Y-%m-%d")),
    x = "Month", y = "Rate (medications per population)",
    colour = "Group", shape = "Group", linetype = "Group"
  ) +
  #theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

post_estimation_trend_plot

# Save as PNG
ggsave(
  filename = "output/results_3_Fig11_post_estimation_trends.png",
  plot = post_estimation_trend_plot,
  path = here::here(),
  width = 10,
  height = 6,
  dpi = 300
)
```
  




##3.Model-checking 

```{r Model-checking }

library(geepack)
library(dplyr)
# ---- 0.  candidate model (baseline for reference) ----------------------------
fit_ar1 <- geeglm(
  medication_count ~ T + X + XT + ZT + ZX + ZXT,
  family  = poisson(link = "log"),
  offset  = log(population),
  id      = practice,
  waves   = T,          # replace with your actual time index column
  corstr  = "ar1",
  data    = practice_level_uti_and_control_long
)

# ------------------------------------------------------------------------------
# 1. QIC — compare working correlation structures
#    (GEE has no likelihood, so no AIC/BIC — use QIC instead)
# ------------------------------------------------------------------------------

fit_indep <- geeglm(
  medication_count ~ T + X + XT + ZT + ZX + ZXT,
  family = poisson(link = "log"),
  offset = log(population),
  id = practice, waves = T,
  corstr = "independence",
  data = practice_level_uti_and_control_long
)

fit_exch <- geeglm(
  medication_count ~ T + X + XT + ZT + ZX + ZXT,
  family = poisson(link = "log"),
  offset = log(population),
  id = practice, waves = T,
  corstr = "exchangeable",
  data = practice_level_uti_and_control_long
)

QIC(fit_ar1)
QIC(fit_exch)
QIC(fit_indep)

# Tabulate side by side
qic_table <- tibble(
  structure = c("ar1", "exchangeable", "independence"),
  QIC       = c(QIC(fit_ar1)["QIC"], QIC(fit_exch)["QIC"], QIC(fit_indep)["QIC"])
)
qic_table %>% arrange(QIC)
# Lower QIC = better trade-off of fit vs. complexity. Pick that structure.
```

# ------------------------------------------------------------------------------
#2. Estimated within-cluster correlation (alpha)
# ------------------------------------------------------------------------------

```{r }
summary(fit_ar1)$corr        # the estimated correlation parameter(s)
# For ar1, one alpha = lag-1 correlation. Near 0 -> independence would suffice.
```

# ------------------------------------------------------------------------------
# 3. Dispersion / overdispersion check
#-------------------------------------------------------------------------------

```{r }
summary(fit_ar1)$dispersion
```

# Manual Pearson-based check (rule of thumb: >1.5-2 = meaningful overdispersion)
```{r }
pearson_resid <- residuals(fit_ar1, type = "pearson")
n_params <- length(coef(fit_ar1))
n_obs    <- length(pearson_resid)
dispersion_manual <- sum(pearson_resid^2) / (n_obs - n_params)
dispersion_manual
```

# ------------------------------------------------------------------------------
# 4. Residual diagnostics
#-------------------------------------------------------------------------------
```{r residual-diagnostics-data}
df_diag <- practice_level_uti_and_control_long %>%
  mutate(
    pearson_resid = residuals(fit_ar1, type = "pearson"),
    fitted_vals   = fitted(fit_ar1)
  )
```

```{r residuals-vs-fitted, fig.width=10, fig.height=6, fig.cap="Figure 12. Residuals vs fitted values from the AR(1) GEE model."}
# Residuals vs fitted (look for trend/funnel = misspecified mean or variance)

# Save as PNG
png(here::here("output", "results_3_Fig12_residuals_vs_fitted.png"), width = 10, height = 6, units = "in", res = 300)
plot(df_diag$fitted_vals, df_diag$pearson_resid,
     xlab = "Fitted values", ylab = "Pearson residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
dev.off()

plot(df_diag$fitted_vals, df_diag$pearson_resid,
     xlab = "Fitted values", ylab = "Pearson residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
```

```{r residuals-vs-time, fig.width=10, fig.height=6, fig.cap="Figure 13. Residuals vs time from the AR(1) GEE model."}
# Residuals vs time

# Save as PNG
png(here::here("output", "results_3_Fig13_residuals_vs_time.png"), width = 10, height = 6, units = "in", res = 300)
plot(df_diag$T, df_diag$pearson_resid,
     xlab = "Time", ylab = "Pearson residuals",
     main = "Residuals vs Time")
abline(h = 0, col = "red", lty = 2)
dev.off()

plot(df_diag$T, df_diag$pearson_resid,
     xlab = "Time", ylab = "Pearson residuals",
     main = "Residuals vs Time")
abline(h = 0, col = "red", lty = 2)
```

```{r residuals-by-practice, fig.width=10, fig.height=6, fig.cap="Figure 14. Pearson residuals over time by practice."}
# Residuals by practice (spot clusters with systematically poor fit)
library(ggplot2)
residuals_by_practice_plot <- ggplot(df_diag, aes(x = T , y = pearson_resid, group = practice)) +
  geom_line(alpha = 0.2) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Pearson residuals over time by practice")

residuals_by_practice_plot

# Save as PNG
ggsave(
  filename = "output/results_3_Fig14_residuals_by_practice.png",
  plot = residuals_by_practice_plot,
  path = here::here(),
  width = 10,
  height = 6,
  dpi = 300
)
```

#-------------------------------------------------------------------------------
# 6. Sensitivity check: 
#         do effect estimates change much across correlation structure choices?
#-------------------------------------------------------------------------------

```{r bind_rows, fig.width=10, fig.height=6, fig.cap="Figure 15. Sensitivity of rate-ratio estimates to the working correlation structure (exponentiated)."}

sensitivity_plot_exp <- bind_rows(
  tidy(fit_ar1,   exponentiate = TRUE, conf.int = TRUE) %>% mutate(structure = "ar1"),
  tidy(fit_exch,  exponentiate = TRUE, conf.int = TRUE) %>% mutate(structure = "exchangeable"),
  tidy(fit_indep, exponentiate = TRUE, conf.int = TRUE) %>% mutate(structure = "independence")
) %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = term, y = estimate, color = structure)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                   position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_flip() +
  labs(y = "Rate ratio (95% CI)", x = NULL,
       title = "Sensitivity of estimates to working correlation structure")

sensitivity_plot_exp

# Save as PNG
ggsave(
  filename = "output/results_3_Fig15_sensitivity_correlation_structure_rr.png",
  plot = sensitivity_plot_exp,
  path = here::here(),
  width = 10,
  height = 6,
  dpi = 300
)
```


#bind_rows_1


```{r bind_rows_1, fig.width=10, fig.height=6, fig.cap="Figure 16. Sensitivity of rate-ratio estimates to the working correlation structure (manually exponentiated)."}

sensitivity_plot_manual_exp <- bind_rows(
  tidy(fit_ar1,   conf.int = TRUE) %>% mutate(structure = "ar1"),
  tidy(fit_exch,  conf.int = TRUE) %>% mutate(structure = "exchangeable"),
  tidy(fit_indep, conf.int = TRUE) %>% mutate(structure = "independence")
) %>%
  mutate(estimate = exp(estimate),
         conf.low  = exp(conf.low),
         conf.high = exp(conf.high)) %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = term, y = estimate, color = structure)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                   position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_flip() +
  labs(y = "Rate ratio (95% CI)", x = NULL,
       title = "Sensitivity of estimates to working correlation structure (manually exponentiated)")

sensitivity_plot_manual_exp

# Save as PNG
ggsave(
  filename = "output/results_3_Fig16_sensitivity_correlation_structure_rr_manual.png",
  plot = sensitivity_plot_manual_exp,
  path = here::here(),
  width = 10,
  height = 6,
  dpi = 300
)
```

