---
title: "WP2P4-Descriptive Analysis: tables and figures"
author: "Arnaud"
date: "2026-07-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(readr)
library(knitr)
library(here)
library(tidyverse)
library(lubridate)
library(scales)
library(dplyr)
```
# Dataset

```{r}
dataset <- read_csv(
  here("output", "dataset_patients_combined.csv.gz"),
  col_types = cols(
    patient_id = col_integer(),
    age = col_double()
  )
)

dataset <- as.data.frame(dataset %>%
  select( 
  patient_id,
  index_date,sex,
  imd,ethnicity,
  stp,region,
  has_medication,numerator_gp_nitrofurantoin_uti,
  numerator_pf_nitrofurantoin_uti
))
dataset$index_date <- as.Date(dataset$index_date)
#head(dataset)
```

# Table 0_0. Nitrofurantoin prescribing volume by region
```{r}
table_0_0 <- dataset %>%
  group_by(region) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),

    gp_nitro_volume = sum(numerator_gp_nitrofurantoin_uti, na.rm = TRUE),
    pf_nitro_volume = sum(numerator_pf_nitrofurantoin_uti, na.rm = TRUE),
    total_nitro_volume = gp_nitro_volume + pf_nitro_volume,

    .groups = "drop"
  ) %>%
  arrange(region)

kable(
  table_0_0,
  caption = "Table 0_0. Nitrofurantoin prescribing volume by region."
)
```
#
```{r fig.width=10, fig.height=5, fig.cap="Figure 0_0. Nitrofurantoin prescribing volume over time by region."}
table_region_plot <- dataset %>%
  group_by(index_date, region) %>%
  summarise(
    gp_nitro_volume = sum(numerator_gp_nitrofurantoin_uti, na.rm = TRUE),
    pf_nitro_volume = sum(numerator_pf_nitrofurantoin_uti, na.rm = TRUE),
    total_nitro_volume = gp_nitro_volume + pf_nitro_volume,
    .groups = "drop"
  ) %>%
  filter(region != "Missing")

nitro_region_plot <- ggplot(
  table_region_plot,
  aes(
    x = index_date,
    y = total_nitro_volume,
    colour = region,
    group = region
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Nitrofurantoin prescribing volume",
    colour = "Region"
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

nitro_region_plot

# Save as PNG
ggsave(
  filename = "output/results_0_0_region_nitrofurantoin_volume.png",
  plot = nitro_region_plot,
  path = here::here(),
  width = 10,
  height = 5,
  dpi = 300
)
```
# Table 0_1. Distribution of patients receiving medication by STP
```{r}
table_stp <- dataset %>%
  group_by(stp) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(stp)

kable(
  table_stp,
  caption = "Table 0_1. Distribution of patients receiving medication by STP."
)
```

# Table 1. Distribution of patients receiving medication by IMD
```{r}
table_imd <- dataset %>%
  group_by(imd) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(imd)

kable(
  table_imd,
  caption = "Table 1. Distribution of patients receiving medication by IMD."
)
```

# Figure 1. Patients receiving medication over time by IMD 

```{r fig.width=10, fig.height=5, fig.cap="Figure 1. Patients receiving medication over time by IMD."}
table_imd_plot <- dataset %>%
  group_by(index_date, imd) %>%
  summarise(
    n_patients = sum(has_medication == 0, na.rm = TRUE), # We will use 1 when using real data
    .groups = "drop"
  ) %>%
  filter(imd != "Missing")

imd_plot <- ggplot(
  table_imd_plot,
  aes(
    x = index_date,
    y = n_patients,
    colour = imd,
    shape = imd,
    group = imd
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "IMD",
    shape = "IMD"
    
  ) +
  scale_x_date(
    date_breaks = "1 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
#
imd_plot 
# Save as PNG 
ggsave(
  filename = "output/results_0_imd_plot.png",
  plot = imd_plot,
  path = here::here(),
  width = 7,
  height = 4,
  dpi = 300
)
```

# Table 2. Distribution of patients receiving medication by Region

```{r}
table_region <- dataset %>%
  group_by(region) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(region)

kable(
  table_region,
  caption = "Table 2. Distribution of patients receiving medication by Region."
)
```

# Figure 2. Patients receiving medication over time by Region

```{r fig.width=10, fig.height=5, fig.cap="Figure 2. Patients receiving medication over time by Region."}
table_region_plot <- dataset %>%
  group_by(index_date, region) %>%
  summarise(
     patients = sum(has_medication == 0, na.rm = TRUE), #n_patients = n(),
    .groups = "drop"
  ) %>%
  filter(region != "Missing")

region_plot <- ggplot(
  table_region_plot,
  aes(
    x = index_date,
    y = patients,
    colour = region,
    group = region
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "Region",
    shape = "Region"
  ) +
  scale_x_date(
    date_breaks = "1 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
#
region_plot
ggsave(
  filename = "output/results_0_region_plot.png",
  plot = region_plot,
  path = here::here(),
  width = 7,
  height = 4,
  dpi = 300
)
```
# Table 3. Distribution of patients receiving medication by Sex

```{r}
table_sex <- dataset %>%
  group_by(sex) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(sex)

kable(
  table_sex,
  caption = "Table 3. Distribution of patients receiving medication by sex."
)
```

# Figure 3. Patients receiving medication over time by Sex

```{r fig.width=10, fig.height=5, fig.cap="Figure 3. Patients receiving medication over time by sex."}

table_sex_plot <- dataset %>%
  group_by(index_date, sex) %>%
  summarise(
    patients = sum(has_medication == 0, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(sex))

sex_plot <- ggplot(
  table_sex_plot,
  aes(
    x = index_date,
    y = patients,
    colour = sex,
    group = sex
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "Sex"
  ) +
  scale_x_date(
    date_breaks = "1 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

sex_plot

ggsave(
  filename = "output/results_0_sex_plot.png",
  plot = sex_plot,
  path = here::here(),
  width = 7,
  height = 4,
  dpi = 300
)
```

# Table 4. Distribution of patients receiving medication by Ethnicity

```{r}
table_ethnicity <- dataset %>%
  group_by(ethnicity) %>%
  summarise(
    n_patients = n(),
    n_with_medication = sum(has_medication == 1, na.rm = TRUE),
    percent_with_medication =
      round(100 * n_with_medication / n_patients, 1),
    .groups = "drop"
  ) %>%
  arrange(ethnicity)

kable(
  table_ethnicity,
  caption = "Table 4. Distribution of patients receiving medication by ethnicity."
)
```

# Figure 4. Patients receiving medication over time by Ethnicity

```{r fig.width=10, fig.height=5, fig.cap="Figure 4. Patients receiving medication over time by ethnicity."}
table_ethnicity_plot <- dataset %>%
  group_by(index_date, ethnicity) %>%
  summarise(
    n_patients = sum(has_medication == 0, na.rm = TRUE), #n_patients = n(),
    .groups = "drop"
  ) %>%
  filter(
    !is.na(ethnicity),
    ethnicity != "Missing"
  )

ethnicity_plot <- ggplot(
  table_ethnicity_plot,
  aes(
    x = index_date,
    y = n_patients,
    colour = ethnicity,
    group = ethnicity
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = "Month",
    y = "Number of patients",
    colour = "Ethnicity"
  ) +
  scale_x_date(
    date_breaks = "1 months",
    date_labels = "%Y-%m"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

ethnicity_plot

ggsave(
  filename = "output/results_0_ethnicity_plot.png",
  plot = ethnicity_plot,
  path = here::here(),
  width = 7,
  height = 4,
  dpi = 300
)
```

# Model 
## Interrupted Time Series (ITS) Analysis Using GEE

```{r gee-model, message=FALSE, warning=FALSE}

# Load required packages
library(readr)
library(here)
library(dplyr)
library(lubridate)
library(ggplot2)
library(geepack)

#--------------------------------------------------
# Load measures dataset
#--------------------------------------------------

measures <- read_csv(
  here("output", "data_patients_measures_Arnaud.csv")
)

# Inspect data
names(measures)
table(measures$measure)

# Keep GP medication outcome only
measures <- measures %>%
  filter(measure == "gp_medication_all_conditions")

head(measures)

#--------------------------------------------------------------------
# ITS variables : covariates for time, intervention, and interaction
#--------------------------------------------------------------------

measures <- measures %>%
  mutate(
    month_date = as.Date(interval_start)
  ) %>%
  group_by(practice) %>%
  arrange(month_date, .by_group = TRUE) %>%
  mutate(
    T = row_number() - 1
  ) %>%
  ungroup() %>%
  mutate(
    X = if_else(month_date >= as.Date("2024-02-01"), 1, 0),
    XT = if_else(
      X == 1,
      T - min(T[X == 1]),
      0
    )
  )

head(measures)

#--------------------------------------------------------------------
# Fit (initial GEE model)
#--------------------------------------------------------------------
gee_model_1 <- geeglm(
  numerator ~ T + X + XT +
    offset(log(denominator)),
  id = practice,
  data = measures,
  family = poisson(link = "log"),
  corstr = "ar1"
)
summary(gee_model_1)
exp(coef(gee_model_1))
#--------------------------------------------------------------------
# Simulate data (for illustration only)
#--------------------------------------------------------------------

set.seed(12345)

measures <- measures %>%
  mutate(

    denominator = round(rnorm(n(), mean = 120, sd = 20)),
    denominator = pmax(denominator, 30),

    prescribing_prob =
      0.65 -
      0.0015 * T -
      0.05 * X -
      0.001 * XT,

    prescribing_prob =
      pmin(
        pmax(prescribing_prob, 0.10),
        0.95
      ),

    numerator = rbinom(
      n(),
      size = denominator,
      prob = prescribing_prob
    ),

    prescribing_rate =
      numerator / denominator
  )

#--------------------------------------------------------------------
# Refit model
#--------------------------------------------------------------------

gee_model_2 <- geeglm(
  numerator ~ T + X + XT +
    offset(log(denominator)),
  id = practice,
  data = measures,
  family = poisson(link = "log"),
  corstr = "ar1"
)

summary(gee_model_2)

#--------------------------------------------------------------------
# Predictions
#--------------------------------------------------------------------

measures$pred <- predict(
  gee_model_2,
  type = "response"
)
#--------------------------------------------------------------------
# Aggregate for plotting
#--------------------------------------------------------------------
plot_data <- measures %>%
  group_by(month_date) %>%
  summarise(
    observed = mean(numerator),
    predicted = mean(pred),
    .groups = "drop"
  )
```
#---------------------------------------------------------------------------
# Plot observed -predicted 
#---------------------------------------------------------------------------
```{r fig.width=10, fig.height=5, fig.cap="Figure ITSA -Observed vs predicted GP prescribing."}
ggplot(plot_data, aes(month_date)) +

  geom_line(
    aes(y = observed,
        colour = "Observed"),
    linewidth = 1.2
  ) +

  geom_line(
    aes(y = predicted,
        colour = "Predicted"),
    linewidth = 1.2
  ) +

  geom_vline(
    xintercept = as.Date("2024-02-01"),
    linetype = "dashed",
    colour = "red"
  ) +

  scale_colour_manual(
    values = c(
      Observed = "black",
      Predicted = "blue"
    )
  ) +

  labs(
    x = "Month",
    y = "GP medication count",
    colour = "",
    title = "Interrupted Time Series using GEE",
    subtitle = "Observed and model-predicted GP prescribing"
  ) +
  theme_bw()
#-----------------------------------------------------------------------------------------------
#  Counterfactual predictions
#-----------------------------------------------------------------------------------------------
measures$fitted <- predict(
  gee_model_2,
  type = "response"
)

cf_data <- measures

cf_data$X <- 0
cf_data$XT <- 0

cf_data$counterfactual <- predict(
  gee_model_2,
  newdata = cf_data,
  type = "response"
)

measures$counterfactual <- cf_data$counterfactual

##
library(dplyr)
#Aggregate to monthly averages
plot_data <- measures %>%
  group_by(month_date) %>%
  summarise(
    observed = mean(numerator),
    fitted = mean(fitted),
    counterfactual = mean(counterfactual),
    .groups = "drop"
  )

#--------------------------------------------------
# Plot counterfactual
#--------------------------------------------------

ggplot(plot_data, aes(month_date)) +

  geom_line(
    aes(
      y = observed,
      colour = "Observed"
    ),
    linewidth = 1.2
  ) +

  geom_line(
    aes(
      y = counterfactual,
      colour = "Counterfactual"
    ),
    linewidth = 1.2,
    linetype = "dashed"
  ) +

  geom_vline(
    xintercept = as.Date("2024-02-01"),
    linetype = "dotted"
  ) +

  scale_colour_manual(
    values = c(
      Observed = "black",
      Counterfactual = "blue"
    )
  ) +

  labs(
    x = "Month",
    y = "GP medication count",
    colour = "",
    title = "Interrupted Time Series Using GEE",
    subtitle = "Counterfactual assumes Pharmacy First had not been introduced"
  ) +

  theme_classic()

```
#Table region
```{r}
library(readr)
dataset_patients_combined<- read_csv(here("output", "dataset_patients_combined.csv.gz"))
names(dataset_patients_combined)
library(dplyr)
str(dataset_patients_combined)
table(dataset_patients_combined$include_patient_otitis_media)
dataset_patients_combined <- dataset_patients_combined %>%
filter(start_date > as.Date("2024-02-01"))%>%
  mutate(
    across(
      c(include_patient_otitis_media,
        include_patient_sinusitis,
        include_patient_sore_throat,
        include_patient_insect_bites,
        include_patient_shingles,
        include_patient_impetigo,
        include_patient_uuti,
        include_patient_overall_eligible
      ),
      ~ as.integer(.x)
    )
  )
table(dataset_patients_combined$include_patient_otitis_media)
#Prescribing rate region
GP_medication_rate_by_region <- dataset_patients_combined %>%
group_by(region) %>%
  summarise(
    GP_uti = sum(numerator_gp_medication_uti, na.rm = TRUE) * 1000 /
      sum(include_patient_uuti, na.rm = TRUE),
    
    GP_sinusitis = sum(numerator_gp_medication_sinusitis, na.rm = TRUE) * 1000 /
      sum(include_patient_sinusitis, na.rm = TRUE),
    
    GP_insectbite = sum(numerator_gp_medication_insectbite, na.rm = TRUE) * 1000 /
      sum(include_patient_insect_bites, na.rm = TRUE),
    
    GP_otitismedia = sum(numerator_gp_medication_otitismedia, na.rm = TRUE) * 1000 /
      sum(include_patient_otitis_media, na.rm = TRUE),
    
    GP_sorethroat = sum(numerator_gp_medication_sorethroat, na.rm = TRUE) * 1000 /
      sum(include_patient_sore_throat, na.rm = TRUE),
    
    GP_shingles = sum(numerator_gp_medication_shingles, na.rm = TRUE) * 1000 /
      sum(include_patient_shingles, na.rm = TRUE),
    
    GP_impetigo = sum(numerator_gp_medication_impetigo, na.rm = TRUE) * 1000 /
      sum(include_patient_impetigo, na.rm = TRUE),
    
    GP_all_conditions = sum(numerator_gp_medication_all_conditions, na.rm = TRUE) * 1000 /
      sum(include_patient_uuti +include_patient_sinusitis +
          include_patient_insect_bites +include_patient_otitis_media +
          include_patient_sore_throat +include_patient_shingles + include_patient_impetigo,
        na.rm = TRUE
      ),
    
    GP_acutebronchitis = sum(numerator_gp_medication_acutebronchitis_control,
      na.rm = TRUE ) * 1000 / sum(include_patient_sore_throat,na.rm = TRUE ),
    GP_conjunctivitis = sum(numerator_gp_medication_conjunctivitisallergic_control,
      na.rm = TRUE ) * 1000 / sum(include_patient_insect_bites, na.rm = TRUE),
    GP_candidiasis = sum(numerator_gp_medication_vulvovaginalcandidiasis_control,
      na.rm = TRUE ) * 1000 / sum(include_patient_uuti,na.rm = TRUE))

GP_medication_rate_by_region
library(dplyr)
library(tidyr)
#
GP_medication_rate_by_region_long <- GP_medication_rate_by_region %>%
  pivot_longer(
    cols = -region,
    names_to = "condition",
    values_to = "medications"
  ) %>%
  mutate(
    condition = recode(
      condition,
      GP_uti = "UTI",
      GP_sinusitis = "Sinusitis",
      GP_insectbite = "Insect bites",
      GP_otitismedia = "Acute otitis media",
      GP_sorethroat = "Sore throat",
      GP_shingles = "Shingles",
      GP_impetigo = "Impetigo",
      GP_all_conditions = "All Pharmacy First conditions",
      GP_acutebronchitis = "Acute bronchitis (control)",
      GP_conjunctivitis = "Allergic conjunctivitis (control)",
      GP_candidiasis = "Vulvovaginal candidiasis (control)"
    )
  )
#
GP_medication_rate_by_region_long
medication_table <- GP_medication_rate_by_region_long %>%
  filter(region !="Missing")%>%
  pivot_wider(
    names_from = region,
    values_from = medications
  )
medication_table
library(gt)
medication_table_gt_region <- medication_table %>%
  gt() %>%
  tab_header(
    title = "GP prescribing rate by region"
  ) %>%
  fmt_number(
    columns = -condition,
    decimals = 0,
    use_seps = TRUE
  ) %>%
  cols_label(
    condition = "Condition"
  ) %>%
  tab_source_note(
    source_note = "Values represent the total number of GP medication prescribing events per 1,000 eligible population per condition per region."
  )
medication_table_gt_region
```

#2.GP_medication_rate_by_IMD
```{r GP_medication_rate_by_IMD}

GP_medication_rate_by_IMD <- dataset_patients_combined %>%
  group_by(imd) %>%
  summarise(
    GP_uti = sum(numerator_gp_medication_uti, na.rm = TRUE) * 1000 /
      sum(include_patient_uuti, na.rm = TRUE),

    GP_sinusitis = sum(numerator_gp_medication_sinusitis, na.rm = TRUE) * 1000 /
      sum(include_patient_sinusitis, na.rm = TRUE),

    GP_insectbite = sum(numerator_gp_medication_insectbite, na.rm = TRUE) * 1000 /
      sum(include_patient_insect_bites, na.rm = TRUE),

    GP_otitismedia = sum(numerator_gp_medication_otitismedia, na.rm = TRUE) * 1000 /
      sum(include_patient_otitis_media, na.rm = TRUE),

    GP_sorethroat = sum(numerator_gp_medication_sorethroat, na.rm = TRUE) * 1000 /
      sum(include_patient_sore_throat, na.rm = TRUE),

    GP_shingles = sum(numerator_gp_medication_shingles, na.rm = TRUE) * 1000 /
      sum(include_patient_shingles, na.rm = TRUE),

    GP_impetigo = sum(numerator_gp_medication_impetigo, na.rm = TRUE) * 1000 /
      sum(include_patient_impetigo, na.rm = TRUE),

    GP_all_conditions =
      sum(numerator_gp_medication_all_conditions, na.rm = TRUE) * 1000 /
      sum(
        include_patient_uuti +
        include_patient_sinusitis +
        include_patient_insect_bites +
        include_patient_otitis_media +
        include_patient_sore_throat +
        include_patient_shingles +
        include_patient_impetigo,
        na.rm = TRUE
      ),

    GP_acutebronchitis =
      sum(numerator_gp_medication_acutebronchitis_control, na.rm = TRUE) * 1000 /
      sum(include_patient_sore_throat, na.rm = TRUE),

    GP_conjunctivitis =
      sum(numerator_gp_medication_conjunctivitisallergic_control, na.rm = TRUE) * 1000 /
      sum(include_patient_insect_bites, na.rm = TRUE),

    GP_candidiasis =
      sum(numerator_gp_medication_vulvovaginalcandidiasis_control, na.rm = TRUE) * 1000 /
      sum(include_patient_uuti, na.rm = TRUE),

    .groups = "drop"
  )

  #long
  GP_medication_rate_by_IMD_long <- GP_medication_rate_by_IMD %>%
  pivot_longer(
    cols = -imd,
    names_to = "condition",
    values_to = "rate"
  ) %>%
  mutate(
    condition = recode(
      condition,
      GP_uti = "UTI",
      GP_sinusitis = "Sinusitis",
      GP_insectbite = "Insect bites",
      GP_otitismedia = "Acute otitis media",
      GP_sorethroat = "Sore throat",
      GP_shingles = "Shingles",
      GP_impetigo = "Impetigo",
      GP_all_conditions = "All Pharmacy First conditions",
      GP_acutebronchitis = "Acute bronchitis (control)",
      GP_conjunctivitis = "Allergic conjunctivitis (control)",
      GP_candidiasis = "Vulvovaginal candidiasis (control)"
    )
  )
  #table
  medication_table_IMD <- GP_medication_rate_by_IMD_long %>%
  filter(!is.na(imd)) %>%
  pivot_wider(
    names_from = imd,
    values_from = rate
  )

  medication_table_gt_IMD <- medication_table_IMD %>%
  gt() %>%
  tab_header(
    title = "GP prescribing rate by IMD quintile"
  ) %>%
  fmt_number(
    columns = -condition,
    decimals = 1,
    use_seps = TRUE
  ) %>%
  cols_label(
    condition = "Condition"
  ) %>%
  tab_source_note(
    source_note =
      "Values represent GP medication prescribing events per 1,000 eligible population by IMD quintile."
  )

medication_table_gt_IMD


```
